---
title: "Introduction to menbayes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to menbayes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(menbayes)
```

# Main Functions

The main functions are:

- `simf1g()`: Simulate genotype counts from an F1 population.
- `simf1g()`: Simulate genotype likelihoods from an F1 population.
- `lrt_men_g4()`: Likelihood ratio test for segregation distortion in F1 populations when the genotypes are known.
- `lrt_men_gl4()`: Likelihood ratio test for segregation distortion in F1 populations when using genotype likelihoods.
- `bayes_men_g4()`: Bayesian test for segregation distortion in F1 populations when the genotypes are known.
- `bayes_men_gl4()`: Bayesian test for segregation distortion in F1 populations when using genotype likelihoods.

# When Null is True

We'll first simulate some data where the null of no segregation distortion is true.
```{r}
set.seed(1)
g1 <- 1
g2 <- 2
alpha <- 1/6
xi1 <- 1/3
xi2 <- 1/3
n <- 20
rd <- 10
x <- simf1g(
  n = n, 
  g1 = g1, 
  g2 = g2, 
  alpha = alpha, 
  xi1 = xi1, 
  xi2 = xi2)
gl <- simf1gl(
  n = n, 
  rd = rd, 
  g1 = g1,
  g2 = g2, 
  alpha = alpha, 
  xi1 = xi1,
  xi2 = xi2)
```

The LRT has a large $p$-value, and the Bayes test has a positive log Bayes factor.

```{r}
lout <- lrt_men_g4(x = x, g1 = g1, g2 = g2)
bout <- bayes_men_g4(x = x, g1 = g1, g2 = g2, chains = 1)
```

```{r}
lout$p_value
bout$lbf
```

```{r}
lout_gl <- lrt_men_gl4(gl = gl, g1 = g1, g2 = g2)
bout_gl <- bayes_men_gl4(gl = gl, g1 = g1, g2 = g2, chains = 1, iter = 1000)
```


```{r}
lout_gl$p_value
bout_gl$lbf
```

# When Alternative is True

We get a very small $p$-value and a very negative log Bayes factor.
```{r}
x <- c(stats::rmultinom(n = 1, size = 20, prob = rep(1/5, 5)))
lout <- lrt_men_g4(x = x, g1 = g1, g2 = g2)
bout <- bayes_men_g4(x = x, g1 = g1, g2 = g2, chains = 1)
lout$p_value
bout$lbf
```
